Figure 3 - ChIP

### 2024-12-02 ms_figure_3_chip.Rmd
render html for Fig3
make Fig3 portrait (11.5x11.6 instead of 11.5x11.1)
add density plot with peak-TSS distances
Co-authored-by: Merle Skribbe <merle.skribbe@gmail.com>
Co-authored-by: Charlotte Soneson <charlottesoneson@gmail.com>
Co-authored-by: Michael Stadler <stadler.michael@gmail.com>

@@

```{r plot-fracBound}
#| fig.width: 5
#| fig.height: 8

@@

```{r Fig3}
#| fig.width: 11.5
#| fig.height: 11.1
#| warning: false

>>>>

# Calculate peak distances to nearest TSS

Using bound peaks for each TF (without "common peaks (ubiquitous)"), we
want to measure the distance of peak mid-points to the nearest TSS, using
negative distances to indicate TF binding upstream of the TSS.

```{r peak-mid-to-tss-dist, fig.width=7, fig.height=5}
tss <- GRanges(seqnames = txannot$seqnames, ranges = IRanges(
    start = txannot$TSS.coordinate, width = 1, names = txannot$tx_id),
    strand = txannot$strand)
peakSel <- peakgr[peakgr$peaktype != "common peaks (ubiquitous)"]
is_enr <- as.matrix(mcols(peakSel)[, grep("^is_enr_in[.]", colnames(mcols(peakSel)))])
colnames(is_enr) <- sub("^is_enr_in.", "", colnames(is_enr))
inclTFs <- colnames(is_enr) != "Untagged" & colSums(is_enr) > 0
peakSelAll <- peakSel[unlist(apply(is_enr[, inclTFs], 2, which))]
d <- distanceToNearest(x = resize(peakSelAll, width = 1, fix = "center"), subject = tss)
isUpstream <- ifelse(strand(tss)[subjectHits(d)] == "+", -1, 1) * sign(start(resize(peakSelAll, width = 1, fix = "center"))[queryHits(d)] - start(tss)[subjectHits(d)]) > 0 
signedDistToTSS <- mcols(d)$distance * (2 * (!isUpstream) - 1)

ggplot(data.frame(d = signedDistToTSS), aes(d)) +
    geom_density() +
    labs(x = "Position of peak mid relative to nearest TSS (bp)",
         y = "Density of peaks",
         caption = "Dashed line: TSS, dotted lines: promoter boundaries") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_vline(xintercept = c(-500, 500), linetype = "dotted", color = main_colors[3]) +
    theme_cowplot()
```

@@

```{r plot-fracBound}
#| fig.width: 8
#| fig.height: 5

@@

```{r Fig3}
#| fig.width: 11.5
#| fig.height: 11.6
#| warning: false
